---
title: "RepDatap2"
author: "Shimon Yannay"
date: "2019 M02 28"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research - Project1

## Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?

### Reading the data


```{r, echo=TRUE}
options(warn=-1)
data <- read.csv("repdata_data_StormData.csv")
```

### Gathering relevant data from all

```{r, echo=TRUE}
sub_data <- data[,c("EVTYPE","FATALITIES","INJURIES", "PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")]
sub_data$PROPDMGEXP <- as.character(sub_data$PROPDMGEXP)
sub_data$CROPDMGEXP <- as.character(sub_data$CROPDMGEXP)
agg_fat <- aggregate(sub_data$FATALITIES,by=list(sub_data$EVTYPE), sum)
agg_inj <- aggregate(sub_data$INJURIES,by=list(sub_data$EVTYPE), sum)
```
### Most fatal events
```{r}
## order by fatalities
agg_fat_ord <- agg_fat[order(agg_fat$x, decreasing = T), ]
## order by injuries
agg_inj_ord <- agg_inj[order(agg_inj$x, decreasing = T), ]
## Head of ordered by fatalities
head_fatal <- head(agg_fat_ord)
names(head_fatal) <- c("Event", "Fatalities")
head_fatal

p1 <- ggplot(data=head_fatal, mapping=aes(x=head_fatal$Event, y=head_fatal$Fatalities))+geom_bar(stat="identity")
print(p1)

## Head of ordered by injuries
head_injury <- head(agg_inj_ord)
names(head_injury) <- c("Event", "Injuries")
head_injury
p1 <- ggplot(data=head_injury, mapping=aes(x=head_injury$Event, y=head_injury$Injuries))+geom_bar(stat="identity")
print(p1)
```
## Across the United States, which types of events have the greatest economic consequences?

```{r}
## calculate the exponent of damage costs
getE <- function(char) {
  options(warn=-1)
    if (char %in% c("h", "H"))
        return(2)
    else if (char %in% c("k", "K"))
        return(3)
    else if (char %in% c("m", "M"))
        return(6)
    else if (char %in% c("b", "B"))
        return(9)
    else if (!is.na(as.numeric(char))) 
        return(as.numeric(char))
    else return(0)
}
## Damaged caused tp property
dmg_prop <- sub_data$PROPDMG * 10**getE(sub_data$PROPDMGEXP)

## Damage caused to crop
dmg_crop <- sub_data$CROPDMG * 10**getE(sub_data$CROPDMGEXP)

## Add data to calculate damage caused to Property and Crop
sub_data$economyDMG <- dmg_prop + dmg_crop

agg_eco <- aggregate(sub_data$economyDMG,by=list(sub_data$EVTYPE), sum)
## order by damage
agg_eco_ord <- agg_eco[order(agg_eco$x, decreasing = T), ]
## Head of ordered by damage costs
head(agg_eco_ord)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
